Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2017-03-06 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
added FH graphs to Ch 8

\subsection{Input Data}

Source term model description ....

The first FDS model that was developed was the simultion of the instrumented gypsum wallboard experiments. The first input needed was the definition of the compuational domain.  The domain of the model had the geometry of a cube with 2.4~m on a side.  The resulting number of computational cells for each case ranged from 27,000 for the 80~mm cells to 1,728,000 for the 20~mm cells.  An isometric view of the model outlined by the domain is shown in Figure      .  The face of the wall is located 1.35~m from the front of the domain.  
The wall was modeled with material properties for gypsum board.
Thermal conductivity = 0.28
Specific Heat = 1.00
Density = 582

@@

Instruments – FDS has devices, which are used to simulate thermocouples and total heat flux gauges.  This devices were positioned  in locations similar to the locations used in the physical experiments, in the plume region and on the wall.   A radiative heat flux gauge was also modelled to provide data at the 0.95 m from the centerline of the burner and 1.0 m above the top of the burner.  
In addition, FDS features devices that can not be implemented such as a device, which can be used to examine the heat release rate per unit length.  In this case, this device was used to determine flame height based on that measure.     
Fire 
The burner was modeled as an obstruction, with the same physical size as the burner.  It was modeled with a steady state heat release rate of 888.89 kW/m$^2$.   The chosen fuel reaction was `methane' since the natural gas used in the laboratory experiments contained approximately 95~\% methane.

\section{Comparison of FDS with Source Fire Results}

In Chapter 2 the fires from the burners were characterized.  In this section the results from FDS simulations of the fire generated by the natural gas burner experiments are compared with the experimental results.  The average plume temperatures at elevations from 0.2~m through 2.2~ above the top face of the burner in 0.2~m increments will be compared.  Then the data generated from the instrumented gypsum wallboard fires are compared with results from FDS simulations.  Four burner positions are examined, beginning 0.61~m from the wall to moving to 0.0~m, where one edge of the burner is in contact with the instrumented gypsum wallboard wall.  First the the plume data will be examined based on the size of the computational grid cells used in the FDS simulations.

\subsection{Grid Size Analysis}

>>>>

\subsubsection{Input Data Source Fire Characterization}

The FDS model for simulating the source fires is very basic, consisting of the definition of the computational domain in space and time, a definition of the burner geometry and the combustion model for the fuel, and specification for measurement points that can be used for analysis temperature, heat flux, and flame height.

The computational domain of the model had the geometry of a cube with 2.4~m on a side.  The resulting number of computational cells for each case ranged from 27,000 for the 80~mm cells to 1,728,000 for the 20~mm cells. For the 80 and 40~mm computational cell scenarios, the cases were run as a single mesh.  For the 20~mm case, the computational domain was divided into 8 uniform meshes.  Each mesh was 2.4~m by 1.2~m and 0.6~m tall.  The time domain for each model was chosen to be 300 seconds, which is similar to the burn time fior each of the fuels.

The burner was modeled as an obstruction with a open top area of 0.09 m$^2$, elevated 0.08~m above the test surface.  This is the same physical size of the natural gas burner used in the experiments.  It was modeled with a steady state heat release rate flux of 888.89 kW/m$^2$, or a heat release rate of 80~kW.   The chosen fuel reaction was `methane' since the natural gas used in the laboratory experiments contained approximately 95~\% methane.  In FDS version 6, the default radiant fraction for the `methane' reaction was set to 0.2.

FDS has devices, which are used to simulate thermocouples and total heat flux gauges.  These devices were positioned in locations similar to the locations used in the source fire experiments.  The thermocouples were positioned in a vertical array, centered on the burner, beginning 0.1~m above the burner surface and spaced every 0.1~m upto 1.2~m above the burner surface.  A total heat flux gauge was also modelled to provide data at the 0.95 m from the centerline of the burner and 1.0 m above the top of the burner.  
In addition, FDS features devices that can not be physically implemented.  One such device, is used to determine the heat release rate per unit length.  This device was used to determine flame height based on that measure.

\subsubsection{Comparison of FDS with Source Fire Results}

In Chapter 2 the fires from the burners were characterized.  In this section the results from FDS simulations of the fire generated by the natural gas burner experiments are compared with the experimental results.  The average plume temperatures at elevations from 0.1~m through 1.2~ above the top face of the burner in 0.1~m increments will be compared.

\subsubsection{Grid Size Analysis}
Figure~\ref{FHNG01_RI} shows the comparison of the centerline gas burner temperatures with three different sets of FDS results.  The only variable in the simulations is the RI. The RIs include three different computational cell sizes; 80~mm (RI=5), 40~mm~(RI=10), 20~mm~(RI=20).  The associated 95~\% confidence limits are also shown the temperature value for each height.  The FDS simulated temperatures based on the course cell size of 80~mm (RI=5) has the best agreement with the temperatures along centerline of the plume.  The two simulations with the higher RI values provide less agreement that the course mesh.  This seems to be counter intuitive to the results expected from an increased RI.  However in each of the FDS models, the results closest to the burner seem to indicate that the model may not be `burning' the appropriate amount of fuel (fuel rich) which results in temperatures significantly lower than measured.  This affect of this fuel rich area above the burner results poorer agreement with increasing RI.  It appears that the the RI=10 and RI=20 models follow the correct trend, however the combuston zone is higher than what was measured.

Another comparison can be made examining the `thermocouple' results from FDS to see if the model of the thermocouple provides improved argeement with the temperatures measured via physical thermocouples.  This comparison is shown in Figure~\ref{FHNG01_RI_TC}.  Again the best agreement along the vertical array of temperatures was provided by the RI=5 model.  Not surprisely the sub model of the thermocouple device shifted the FDS values slightly.  Variables such as density and emissivity required for modeling the thermal response and heat transfer to the thermocouple bean are influencing the diffrences between the gas temperature and the mesuared temperature FDS values.

Mc caffery plume

\begin{figure}[h]
	\includegraphics[width=5in]{../Figures/FHNG01_RI} \\

	\caption[Comparison of burner centerline gas temperatures with FDS simulated gas temperatures.]{Comparison of burner centerline gas temperatures with FDS simulated gas temperatures generated with three different RIs 5, 10 and 20.}
	\label{FHNG01_RI}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=5in]{../Figures/FHNG01_RI_TC} \\

	\caption[Comparison of burner centerline gas temperatures with FDS simulated gas temperatures.]{Comparison of burner centerline gas temperatures with FDS simulated gas temperatures generated with three different RIs 5, 10 and 20.}
	\label{FHNG01_RI_TC}
\end{figure}

\subsection{Natural Gas Instrumented Wall Fires}

that was developed was the simultion of the instrumented gypsum wallboard experiments. The first input needed was the definition of the compuational domain.  The domain of the model had the geometry of a cube with 2.4~m on a side.  The resulting number of computational cells for each case ranged from 27,000 for the 80~mm cells to 1,728,000 for the 20~mm cells.  An isometric view of the model outlined by the domain is shown in Figure      .  The face of the wall is located 1.35~m from the front of the domain.  
The wall was modeled with material properties for gypsum board.
Thermal conductivity = 0.28
Specific Heat = 1.00
Density = 582

@@

Instruments – FDS has devices, which are used to simulate thermocouples and total heat flux gauges.  This devices were positioned  in locations similar to the locations used in the physical experiments, in the plume region and on the wall.   A radiative heat flux gauge was also modelled to provide data at the 0.95 m from the centerline of the burner and 1.0 m above the top of the burner.  
In addition, FDS features devices that can not be implemented such as a device, which can be used to examine the heat release rate per unit length.  In this case, this device was used to determine flame height based on that measure.     
Fire

Then the data generated from the instrumented gypsum wallboard fires are compared with results from FDS simulations.  Four burner positions are examined, beginning 0.61~m from the wall to moving to 0.0~m, where one edge of the burner is in contact with the instrumented gypsum wallboard wall.
